---
title: "Basic Staffs"
output: html_notebook
---

## Define Gross Return, Net Return and Log Return

Given a prices vector, calculate the above returns

```{r}
prices <- c(1.3, 1.2)

# Gross return
# Formula Rg(t) = S(t)/S(t-1)
gross <- prices[2] / prices[1]
gross

# Net return
# Formula r(t) = S(t)/S(t-1) - 1
net <- prices[2]/ prices[1] - 1
net

# Formula:
# Rlog(t) = log(1 + r(1)) = log(S(t) / S(t-1)) = log(S(t)) - log(S(t-1))
# You may need to know the logarithm rules (product rule, power rule, quotient rule, etc), see: https://www.google.com.hk/search?q=logarithm+rules&oq=logarithm+rules&aqs=chrome..69i57j0l5.4676j0j7&sourceid=chrome&ie=UTF-8
diff(log(prices))
```

The reason that we use Log Return: 
* A linear price scale is plotted so that there is an equal distance between the prices；
* A logarithmic price scala is plotted in such way that two equal percent changes are plotted as the same vertical distance;

See: http://www.investopedia.com/ask/answers/05/logvslinear.asp

Suppose your earning increase by 10 times every year:

```{r}
earnings = c(1, 10, 100, 1000, 10000)
df <- data.frame(earnings)
df$log_earnings <- log10(earnings)

par(mar = c(5,5,2,5))
with(df, plot(earnings, type='l', col='red3',
              ylab='earning'))
par(new=T)
with(df, plot(log_earnings, type='l', axes=F, xlab=NA, ylab=NA, col='blue3'))
axis(side=4)
mtext(side=4, line=3, 'magnification')
legend("topleft", legend = c('earnings', 'magnification'),
       col=c('red3', 'blue3'),
       lty=c(1,1))
```

## Mean, Variance, Standard Deviation

Mean formula: 
$$\sum_{i=1}^n xp(x)=\mu$$

```{r}
# Mean is also know as the `expected value`
# E(X) = `the weighted average of the values in the range of X`

prob_dice <- rep(1/6, 6) # The probability of rollin a dice (0.1666667 0.1666667 0.1666667 0.1666667 0.1666667 0.1666667)
sum(prob_dice) # should be 100%

dice <- 1:6 # Possible result of rolling a dice
samples <- table(sample(dice, size=10000, prob=prob_dice, replace=T)) # Roll a dice 100 times, and build an contingency table of counts

labels <- sprintf("%0.1f%%", samples / sum(samples) * 100)
bar_samples <- barplot(samples, ylim=c(0, 2000))
text(x=bar_samples, y=samples+10, label=labels)
expected_value <- sum(prob_dice * dice) # 3.5
```

## How to calculate cumulative return of client portfolio

```{r}
library(pacman)
p_load(tidyverse)

dates <- c('2023-07-03', '2023-07-04', '2023-07-05', '2023-07-06', '2023-07-07')
cash_transactions <- c(1000, 0, 2540, -1540, 0)
asset_market_value <- c(1000, 998.23, 3552.63, 2015.42, 2053.87)


df <- tibble(date = dates, 
             cash_tx = cash_transactions, 
             market_val = asset_market_value)

log_ret <- diff(log(asset_market_value - cash_transactions))
log_ret
cum_log_ret <- cumsum(log_ret)
exp(cum_log_ret) - 1

df <- df %>% 
  mutate(market_val_adj = market_val - cash_tx) %>% 
  mutate(log_ret = log(market_val_adj) - log(lag(market_val, 1))) %>%
  mutate(log_ret = replace_na(log_ret, 0)) %>%
  mutate(cum_log_ret = cumsum(log_ret),
         cum_ret = exp(cum_log_ret) - 1)
           
df
```
