---
title: "Time Series Analysis"
output: html_notebook
---

## Rudimentary Knowledge

* ADF test (Augmented Dickey-Fuller Test): `Null Hypothesis` the time ordered set of observations is non-stationary
* Seasonal means model: expected average temperature for each of the 12 months is given as:
** μt = β1 for (1, 13, 25..) β2 for (2, 14, 26..) β12 for (12, 24, 36)
* `~` operation in R is a formula, left of the ~ is the response, right are the explanatory variables
* Regression is methodologies to predict a response variable
* OLS (ordinary least squares): Linear Regression is one type of OLS regression. In OLS regression, a quantitative dependent variable is predicted from a weighted sum of predictor variables. See `R in Action by Robert I. Kabacoff` for equation
* Box-Cox procedure: produce normal distribuated response variable by, e.g. log, 1/(y), sqrt(y)
* acf: autocorrelation function (if all lag > 0 coeffecients are within the confidence interval range, then it is acceptable)

## Regression Test on Whether `tempdub` Dataset is Stationary 

```{r}
library(TSA)
data(tempdub)
plot(tempdub, col='blue')
adf.test(tempdub) # p-value: 0.01, rejected non-stationary null hypothesis

month <- season(tempdub) # Create repeated rows with (Jan, Feb ... Dec)
summary(lm(tempdub ~ month - 1))
# The model accounts 99.57 percent of all variation
```

## `hare` Dataset is Less Stable Because of More Factors

```{r}
library(TSA)
data(hare)
plot(hare, col='blue')
adf.test(hare) # p-value: 0.06, reject stationary alternative hypothesis

# Use Box-Cox procedure to make it stationary
par(mfrow=c(2, 2))
BoxCox.ar(hare) 
# Output of Box-Cox will show that power transform parameter = 0.5
# According to Power Transformation Table; http://webcache.googleusercontent.com/search?q=cache:zQPZyzK3Z7UJ:www.pt.ntu.edu.tw/hmchai/SAS/SASdescriptive/SASnormality.htm+&cd=8&hl=en&ct=clnk&gl=hk
# sqrt transformation is proper
plot(sqrt(hare), col='blue')
acf(sqrt(hare))
acf(hare)
pacf(sqrt(hare))
```

