---
title: "Visualize Securities Datasets"
output: html_notebook
---

## Hige dimensional undirected graph estimation package (huge) has an S&P 500 stocks dataset

```{r}
library(huge)
data(stockdata)
attach(stockdata)

# Each column represents a stock
D <- length(data[1,])

# Each row contains the prices of a specific time
len <- length(data[,1])
prices <- data[,1:D]

# First column is stock symbol
lab <- stockdata$info[1:D,1]

display_charts <- function(prices, lab, nrow=3, ncol=4, sleep_secs=4) {
  dims <- length(prices[1,])
  for(chart_grp in (1:ceiling(dims/(nrow*ncol)))) {
    print(chart_grp)
    par(mar=c(3.82, 1.82, 1.82, 0.82))
    par(mfrow=c(nrow, ncol))
    for(i in 1:(nrow*ncol)) {
      # j is the jth column in prices
      j <- ((chart_grp-1)*nrow*ncol+i)
      # if j is with total d
      if(j <= dims) {
        print(paste(j, lab[j]))
        plot(prices[,j], type='l', xlab=paste(j, lab[j]))
      }
    }
    Sys.sleep(sleep_secs)
  }
}

# Unit test
display_charts(prices, lab)
```

## Adjusting for Stock Splits

* When a stock's price has appreciated greatly, splitting the common stock makes the price more affordable
* Also, it can enhance visibility into the Company's profitability on a per-share basis (e.g. JDSU on Sep, 2016)

```{r}
# Rules to adjust according prices split signals, 
split_adjust <- function(prices, symbol) {
  len = length(prices)
  # The last price tick
  original_final_price <- prices[len]
  # j is the second of the pair for comparsion, so it starts from 2
  for(j in 2:len) {
    split <- 0  # Initial split is 0
    
    # Rules to determine whether a stock's price is splitted
    if (prices[j-1] >= 1.4*prices[j]) {
      split = 1.5  # 3 for 2 use 3 shares in exchanged for 2 shares
      if (prices[j-1] >= 1.8*prices[j]) split = 2    # 2 for 1
      if (prices[j-1] >= 2.9*prices[j]) split = 3    # 3 for 1
      if (prices[j-1] >= 3.9*prices[j]) split = 4    # 4 for 1
      if (prices[j=1] >= 4.9*prices[j]) stop(paste(symbol, 'detected more than 4:1 split'))
      print(paste('Split adjusting', symbol, j, split, prices[j-1], prices[j]))
    }
    
    # Rules to determine whether a stock's price is reverse splitted
    if (prices[j-1] <= prices[j]/1.4) {
      split = -1.5 
      if (prices[j-1] <= prices[j]/1.9 && prices[j-1] >= prices[j]/2.1) split = -2
      if (prices[j-1] <= prices[j]/2.9 && prices[j-1] >= prices[j]/3.1) split = -3
      if (prices[j-1] <= prices[j]/5.8 && prices[j-1] >= prices[j]/6.2) split = -6
      if (prices[j-1] <= prices[j]/7.7 && prices[j-1] >= prices[j]/8.3) split = -8
      if (prices[j-1] <= prices[j]/9.7 && prices[j-1] >= prices[j]/10.3) split = -10
      if ((split == 0) && (prices[j-1] <= prices[j]/2.9)) stop(paste(symbol, 'detected more than double reverse split'))
      print(paste('Reverse split adjusting', symbol, j, split, prices[j-1], prices[j])) 
    }
    
    if (split != 0) {
      # If there is a split, apply the split to prices vector
      # `j` is the where in a price vector the split incurs
      for (k in j:len) {
       if (split > 0) prices[k] = prices[k] * split
       else prices[k] = prices[k] / abs(split)
      }
    }
  }
  
  final_price = prices[len]
  
  # Before this step, prices before splits will be kept
  # After this step, prices before last split will be adjusted
  return(prices * original_final_price / final_price)
}

# Unit test
p <- c(3.0, 3.0, 2.0, 11.88, 5.9, 1.95, 3.9, 3.9, 1.5, .75, 1.00, 1.2, 1.4, 1.8, 2.1, 1.05, 1.3)
sap <- split_adjust(p, "SYM")
plot(p, type='l', ylim=c(0, 15)); points(sap, col=4)

# Apply adjust split to stock 'JDSU'
library(huge)
data(stockdata)
attach(stockdata)
# Each column represents a stock
total_stocks_num <- length(data[1,])
# Each row contains the prices of a specific time
len <- length(data[,1])
prices <- data[, 1:total_stocks_num]
# First column is stock symbol
lab <- stockdata$info[1:total_stocks_num, 1]

idx <- match('JDSU', lab)
sap <- split_adjust(prices[, idx], 'JDSU')
plot(prices[, idx], type='l', ylim=c(0, 55), xlab='JDSU'); points(sap, col=4)
```

## Find log returns for the prices series

```{r}
find_returns <- function(prices, split_adjusted=T) {
  len <- dim(prices)[1]
  total_stocks_num <- dim(prices)[2]
  log_returns <- matrix(nrow=(len-1), ncol=total_stocks_num)
  for (i in 1:total_stocks_num) {
    if (!split_adjusted) prices[, i] <<- split_adjust(prices[, i], lab[i])
    print(diff(log(prices[, i])))
    log_returns[, i] <- 100 * diff(log(prices[, i]))
  }
  return(log_returns)
}

returns <- find_returns(prices, split_adjusted=F)
returns
```

## Plotting mulitple series data in 1 chart

This chart scales all prices to 1 unit of curreny. This scaling replaces the actual prices with relative prices so that securities returns can be compared.

```{r}
plot_multi_series <- function(prices, lab, w, total_stocks_num, cc='days', ret=NA, ylim=c(.2, 15), is_alone=T) {
  if (is_alone) plot.new()
  
  # Each stock in the same chart shall have a different color
  map_color <- function(d) {
    if (d %% 8 == 7) 1 
    else if (d == 8) 2
    else if (d == 15) 3
    else if (d == 23) 4
    else d
  }
  
  # Set the margin of the graphical parameters
  par(mar=c(4, 2.82, 1.82, 1))
  if (is_alone) par(mfrow=c(1, 1))
  
  total <- 0; len <- dim(prices)[1]
  first <- T; total_stocks_num <- dim(prices)[2]
  
  for (d in 1:total_stocks_num) {
    # Make sure the first prices value of the time series is not NA
    if(!is.na(prices[1, d]) && !is.na(w[d]) && w[d] > 0) {
      print(lab[d])
      total <- total + 1
      if (first) {
        # If this is the first plot, use plot function
        first = F
        # Every prices divided by prices of the first day, line chart, xlab is displayed in days
        plot(prices[, d]/prices[1, d], type='l', col=map_color(d), xlab=cc, ylim=ylim)
      } else {
        lines(prices[, d]/prices[1, d], type='l', col=map_color(d))
      }
      # cex: character expansion factor to determine font size
      # len: display the texts at the end of X-axis
      # prices[len, d]/prices[1, d]: height of the text
      text(len, (prices[len, d]/prices[1, d]), lab[d], col=map_color(d), cex=.8)
    }
  }
}

# Unit test
stocks_num_to_draw <- 12
w <- rep(1/stocks_num_to_draw, stocks_num_to_draw)
plot_multi_series(prices, lab, w, stocks_num_to_draw, cc=paste(sum(w > 0), 'stocks'), ret='', ylim=c(.5, 8))
```

